---
title: "R Notebook"
output: html_notebook
---
# Group 5, Matthew Hines, Emmanuel Garzo, Dustin Beaver

# Guiding Question

What factors are common between all hall of fame MLB baseball players?
  
# Become Acquainted

* Data.World
* Authors of the data set are 
    * Chris Dalzell
    * Michael Friendly
    * Dennis Murphys
    * Martin Monkman
    * Vanessa Foot & Justeena Zaki-Azat [vignettes]
* Collected orginally in 1996, there is no given reason for collection.
* There are 25 data sets included. the main focus will be on HallofFame, Pitching, Hitting, Fielding and Teams.
* The varaibles that will be mainly used are PLayerID, which is unique to each player and allow us to join tables and track player statistics 

# Explore Intuition Related to the Research Question 

```{r}
# Number of inducted versus not inducted per year
 ggplot( data = HallOfFame, aes(x = yearID)) + geom_histogram(binwidth = 3.3) + facet_wrap(~inducted, ncol = 4) + labs(title = "") 
```
```{r}
# Home Runs per Year
ggplot( data = Batting, aes(x = yearID)) + geom_histogram(binwidth = 5.9) + labs(title = "") 
```
# TWO Data Sources

* Lahman's Baseball DataBase, will be sourced by CSV files
* RetroSheet Game Log Data Set, it's also a CSV 
