---
title: "R Notebook"
author: "Group 5, Matthew Hines, Emmanuel Garzo, Dustin Beaver"
date: "4/5/2020"
output: html_notebook
---


# Guiding Question

What factors are common between all hall of fame MLB baseball players?
  
# Become Acquainted

* Data.World
* Authors of the data set are 
    * Chris Dalzell
    * Michael Friendly
    * Dennis Murphys
    * Martin Monkman
    * Vanessa Foot & Justeena Zaki-Azat [vignettes]
* Collected orginally in 1996, there is no given reason for collection.
* There are 25 data sets included. the main focus will be on HallofFame, Pitching, Hitting, Fielding and Teams.
* The varaibles that will be mainly used are PLayerID, which is unique to each player and allow us to join tables and track player statistics 

# Explore Intuition Related to the Research Question 
```{r}
# Number of inducted versus not inducted per year
 ggplot( data = HallOfFame, aes(x = yearID)) + geom_histogram(binwidth = 3.3) + facet_wrap(~inducted, ncol = 4) + labs(title = "") 
```
This plot shows the number of inducted players versus not inducted per year.

```{r}
# Home Runs per Year
ggplot( data = Batting, aes(x = yearID)) + geom_histogram(binwidth = 5.9) + labs(title = "") 
```
This plot shows the number of home runs per year. 

### Analyzing the Data
While analyzing the data, we noticed that having an excellent batting average gives you a better chance to be in the hall of fame.
We saw some potential outliers because pitchers dont have a good batting average but they also could be in the hall of fame.

# TWO Data Sources

* Lahman's Baseball DataBase, will be sourced by CSV files
* RetroSheet Game Log Data Set, it's also a CSV 
